---
title: "Basic Usage of the Boilerplate Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Usage of the Boilerplate Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(boilerplate)
```

# Introduction

The `boilerplate` package helps you manage and generate standardized text for scientific reports, particularly for methods and results sections. It supports template variable substitution, hierarchical organization of text, and provides tools for creating structured, reusable content.

This vignette covers the basic usage of the package:

1. Managing text databases
2. Managing measures databases
3. Generating text with variable substitution
4. Creating complete methods and results sections

## 1. Managing Text Databases

First, let's initialize the default databases:

```{r init}
# Initialize default databases (do this once per project)
boilerplate_init_text()
```

Now we can view the default methods database:

```{r list_methods}
# List the methods database
init_methods_db <- boilerplate_manage_text(category = "methods", action = "list")

# See what's available at the top level
names(init_methods_db)
```

### Adding New Text

Let's add a new method entry:

```{r add_method}
# Add a new method entry
methods_db <- boilerplate_manage_text(
  category = "methods",
  action = "add",
  name = "sample_selection",
  value = "Participants were selected from {{population}} during {{timeframe}} using {{sampling_method}} sampling.",
  db = init_methods_db
)

cat(methods_db$sample_selection)
```

### Nested Entries

The package supports hierarchical organization using dot notation:

```{r add_nested}
# Add a nested entry
methods_db <- boilerplate_manage_text(
  category = "methods",
  action = "add",
  name = "statistical.machine_learning.random_forest",
  value = "We used random forests with {{n_trees}} trees and {{max_depth}} maximum depth.",
  db = methods_db
)

# View the nested structure
statistical_methods <- boilerplate_manage_text(
  category = "methods",
  action = "list",
  name = "statistical",
  db = methods_db
)
str(statistical_methods, max.level = 2)
```

### Retrieving Text with Variables

We can retrieve text with template variable substitution:

```{r get_text}
# Get text with variables substituted
sample_text <- boilerplate_manage_text(
  category = "methods",
  action = "get",
  name = "sample_selection",
  template_vars = list(
    population = "university students",
    timeframe = "2020-2021",
    sampling_method = "convenience"
  ),
  db = methods_db
)

cat(sample_text)
```

## 2. Managing Measures Databases

The package provides special functions for managing measures. 

First First, initialise the default databases:

```{r init}
# Initialize default databases (do this once per project)
boilerplate_init_measures()
```

```{r measures}
# List the measures database
measures_db <- boilerplate_manage_measures(action = "list")

# Add a new measure
measures_db <- boilerplate_manage_measures(
  action = "add",
  name = "anxiety.gad7",
  measure = list(
    description = "Anxiety was measured using the GAD-7 scale.",
    reference = "spitzer2006",
    waves = "1-3",
    keywords = c("anxiety", "mental health", "gad"),
    items = list(
      "Feeling nervous, anxious, or on edge",
      "Not being able to stop or control worrying",
      "Worrying too much about different things",
      "Trouble relaxing"
    )
  ),
  db = measures_db
)

# Retrieve the measure
gad7 <- boilerplate_manage_measures(
  action = "get",
  name = "anxiety.gad7",
  db = measures_db
)

str(gad7)



```

## 3. Generating Text with Variable Substitution

The `boilerplate_generate_text()` function allows us to combine multiple text sections with variable substitution:

```{r generate_text}
# Generate text combining multiple sections
methods_text <- boilerplate_generate_text(
  category = "methods",
  sections = c("sample_selection", "statistical.machine_learning.random_forest"),
  global_vars = list(
    population = "university students",
    timeframe = "2020-2021",
    sampling_method = "convenience",
    n_trees = 500,
    max_depth = 10
  ),
  db = methods_db,
  add_headings = TRUE,
  heading_level = "##"
)

cat(methods_text)

# define variables 


path_measures_test = '/Users/joseph/GIT/templates/databases/measures'

test_measures_db <- margot::here_read("merged_db", path_measures_test)
exposure_var = "perfectionism"
outcome_vars = c("belong", "agreeableness")


# Generate measures text
measures_text <- boilerplate_measures_text(
  exposure_var = exposure_var,
  outcome_vars = outcome_vars,
  measures_db = test_measures_db,  # For measure information
  appendices_measures = "Appendix A",
  print_waves = TRUE)

cat(measures_text)

  
  str(test_measures_db)
boilerplate_measures_text(
  variable_heading = "Exposure Variable",
  variables = exposure_var,
  measures_db = test_measures_db,  # For measure information
  appendices_measures = "Appendix A",
  print_waves = TRUE)

```

## 4. Creating Complete Methods and Results Sections

The package provides specialized functions for generating methods and results sections:

```{r methods_results}
# Define study parameters
study_params <- list(
  exposure_var = "treatment",
  population = "university students",
  timeframe = "2020-2021",
  effect_size = "0.35",
  confidence_interval = "95% CI: 0.21, 0.49",
  interpretation = "a moderate positive effect"
)

# Generate methods text with appropriate defaults
methods_text <- boilerplate_methods_text(
  sections = c("sample", "causal_assumptions.identification"),
  global_vars = study_params,
  db = methods_db,
  add_headings = TRUE
)

cat(methods_text)

# Generate results text
results_text <- boilerplate_results_text(
  sections = c("main_effect"),
  results_data = study_params,
  add_headings = TRUE
)

cat(results_text)
```

### Generating Measures Text for Reporting

For reporting measures used in a study:

```{r measures_text}
# Define measures used in the study
measures_db <- boilerplate_add_measure(
  name = "depression.phq9",
  description = "Depression was measured using the PHQ-9 scale.",
  reference = "kroenke2001",
  waves = "1-3",
  items = c("Little interest or pleasure in doing things", 
            "Feeling down, depressed, or hopeless"),
  keywords = c("depression", "mental health"),
  db = measures_db
)

# Define the exposure and outcome variables
exposure_var <- "treatment"
outcome_vars <- list(
  psychological = c("anxiety.gad7", "depression.phq9")
)

# Generate measures text
measures_text <- boilerplate_measures_text(
  exposure_var = exposure_var,
  outcome_vars = outcome_vars,
  measures_db = measures_db,
  appendices_measures = "Appendix A",
  print_waves = TRUE,
  db = measures_db
)

cat(measures_text)
```

## 5. Saving Databases for Future Use

To save your customized databases for future use:

```{r save_db, eval=FALSE}
# Save the methods database
boilerplate_manage_text(
  category = "methods",
  action = "save",
  db = methods_db
)

# Save the measures database
boilerplate_manage_measures(
  action = "save",
  db = measures_db
)
```

## 6. Simplified Wrapper Functions

The package provides simplified wrapper functions for common operations:

```{r wrappers}
# Add text with a simplified function
methods_db <- boilerplate_add_text(
  category = "methods",
  name = "data_analysis",
  text = "Data were analyzed using {{software}} version {{version}}.",
  db = methods_db
)

# Get text with a simplified function
analysis_text <- boilerplate_get_text(
  category = "methods",
  name = "data_analysis",
  vars = list(software = "R", version = "4.2.0"),
  db = methods_db
)

cat(analysis_text)

# Add a measure with a simplified function
measures_db <- boilerplate_add_measure(
  name = "wellbeing",
  description = "Wellbeing was measured using a 5-item scale.",
  reference = "smith2020",
  items = c("I am satisfied with my life", "My life has meaning"),
  keywords = c("wellbeing", "satisfaction"),
  db = measures_db
)
```

# Conclusion

The `boilerplate` package provides a flexible framework for managing and generating standardized text for scientific reports. By creating reusable text templates and organizing them hierarchically, you can maintain consistency across reports while easily customizing content with template variables.

For more advanced usage, see the "Research Workflow" vignette.
