---
title: "Catppuccin Theme Example"
subtitle: "A cosy presentation template"
author: "Joseph A. Bulbulia"
date: today
format:
  revealjs:
    theme: dark
    css: /Users/joseph/GIT/templates/css/catppuccin-mocha.css
    slide-number: true
    preview-links: auto
    code-fold: true
    code-overflow: scroll
    code-line-numbers: true
    highlight-style: monokai
    progress: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tidyr)
source("/Users/joseph/GIT/templates/R/theme_catppuccin.R")
```

::: {.notes}
welcome everyone. this talk demonstrates the catppuccin mocha theme for quarto revealjs presentations.
:::

## overview

this example demonstrates the **catppuccin mocha** theme for quarto revealjs presentations.

- warm, pastel colours
- easy on the eyes
- consistent aesthetic

::: {.notes}
speaker notes appear here. press 's' during the presentation to open the speaker view. you'll see the current slide, next slide, elapsed time, and these notes.
:::

## code example

```{r}
#| echo: true
#| eval: false

# fit a simple linear model
library(ggplot2)

model <- lm(mpg ~ wt + hp, data = mtcars)
summary(model)
```

::: {.notes}
this slide demonstrates syntax highlighting for R code. the code-fold option in the YAML lets you collapse code blocks.
:::

## ggplot2: scatter plot

```{r}
#| echo: true
#| fig-width: 9
#| fig-height: 4.5

source("/Users/joseph/GIT/templates/R/theme_catppuccin.R")

ggplot(mtcars, aes(x = wt, y = mpg, colour = factor(cyl))) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_colour_ctp() +
  labs(
    title = "fuel efficiency by weight",
    x = "weight (1000 lbs)",
    y = "miles per gallon",
    colour = "cylinders"
  ) +
  theme_catppuccin()
```

::: {.notes}
this scatter plot uses theme_catppuccin() and scale_colour_ctp() for consistent styling.
:::

## ggplot2: bar chart

```{r}
#| echo: true
#| fig-width: 9
#| fig-height: 4.5

source("/Users/joseph/GIT/templates/R/theme_catppuccin.R")

ggplot(mpg, aes(x = class, fill = class)) +
  geom_bar() +
  scale_fill_ctp() +
  labs(
    title = "vehicle classes in mpg dataset",
    x = "class",
    y = "count"
  ) +
  theme_catppuccin() +
  theme(legend.position = "none")
```

::: {.notes}
bar chart using scale_fill_ctp() for the fill colours. legend hidden as colours are self-explanatory.
:::

## ggplot2: boxplot

```{r}
#| echo: true
#| fig-width: 9
#| fig-height: 4.5

source("/Users/joseph/GIT/templates/R/theme_catppuccin.R")

ggplot(iris, aes(x = Species, y = Sepal.Length, fill = Species)) +
  geom_boxplot(alpha = 0.8, colour = ctp_mocha$text) +
  scale_fill_ctp() +
  labs(
    title = "sepal length by species",
    x = "species",
    y = "sepal length (cm)"
  ) +
  theme_catppuccin() +
  theme(legend.position = "none")
```

::: {.notes}
boxplot with catppuccin fill colours. outline uses the text colour for visibility.
:::

## ggplot2: density plot

```{r}
#| echo: true
#| fig-width: 9
#| fig-height: 4.5

source("/Users/joseph/GIT/templates/R/theme_catppuccin.R")

ggplot(iris, aes(x = Sepal.Width, fill = Species)) +
  geom_density(alpha = 0.6) +
  scale_fill_ctp() +
  labs(
    title = "sepal width distribution",
    x = "sepal width (cm)",
    y = "density"
  ) +
  theme_catppuccin()
```

::: {.notes}
overlapping density plots with transparency. the catppuccin palette maintains readability even with overlap.
:::

## ggplot2: faceted plot

```{r}
#| echo: true
#| fig-width: 9
#| fig-height: 4.5

source("/Users/joseph/GIT/templates/R/theme_catppuccin.R")

ggplot(mpg, aes(x = displ, y = hwy, colour = factor(cyl))) +
  geom_point(alpha = 0.7) +
  facet_wrap(~drv) +
  scale_colour_ctp() +
  labs(
    title = "highway mpg by displacement and drive type",
    x = "displacement (L)",
    y = "highway mpg",
    colour = "cylinders"
  ) +
  theme_catppuccin()
```

::: {.notes}
faceted plot showing how the theme handles strip backgrounds and multiple panels.
:::

## ggplot2: line chart

```{r}
#| echo: true
#| fig-width: 9
#| fig-height: 4.5

source("/Users/joseph/GIT/templates/R/theme_catppuccin.R")

economics_long <- tidyr::pivot_longer(
  economics[, c("date", "unemploy", "pce")],
  cols = -date,
  names_to = "variable",
  values_to = "value"
)

ggplot(economics_long, aes(x = date, y = value, colour = variable)) +
  geom_line(linewidth = 1) +
  scale_colour_ctp() +
  labs(
    title = "us economic indicators over time",
    x = "date",
    y = "value",
    colour = "indicator"
  ) +
  theme_catppuccin()
```

::: {.notes}
time series line chart. colours remain distinct and readable on the dark background.
:::

## maths support

inline: $\hat{\beta} = (X'X)^{-1}X'y$

display:

$$
\tau_i(x) = E[Y_i(1) - Y_i(0) \mid X_i = x]
$$

::: {.notes}
- the first equation is the OLS estimator
- the second is the conditional average treatment effect (CATE)
- mathjax renders these beautifully
:::

## emphasis styles

- **bold text** stands out in peach
- *italic text* appears in rosewater
- `inline code` uses green on surface

> blockquotes have a lavender accent

::: {.notes}
catppuccin uses semantic colours: peach for emphasis, rosewater for subtle highlights, green for code. the palette is designed for readability.
:::

## tables

| outcome | estimate | 95% CI |
|---------|----------|--------|
| wellbeing | 0.23 | [0.12, 0.34] |
| cooperation | 0.18 | [0.08, 0.28] |
| meaning | 0.31 | [0.19, 0.43] |

::: {.notes}
tables use alternating row colours (surface0 and mantle) for readability. headers are styled with lavender text on surface1 background.
:::

## callouts

::: {.callout-note}
this is a note callout
:::

::: {.callout-tip}
this is a tip callout
:::

::: {.callout-warning}
this is a warning callout
:::

::: {.notes}
callouts use catppuccin accent colours: blue for notes, green for tips, yellow for warnings, red for important/caution.
:::

## links and references

- [catppuccin palette](https://github.com/catppuccin/catppuccin)
- [quarto revealjs docs](https://quarto.org/docs/presentations/revealjs/)

::: {.notes}
links use sapphire (#74c7ec) and hover to sky (#89dceb). these are easy to distinguish from regular text.
:::

## thank you

questions?

::: {.notes}
remember to:

- thank the organisers
- mention collaborators
- point to the github repo for code
:::

